---
title: "Coffee Clustering"
author: "Jeremy Leipzig"
date: "April 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 8.10 Estimating coffee competition in market segments

**What measure would you use to reflect the level of competition between different instant coffee alternatives?**
Consumers who rotate between the same two or three coffees should form market segments. To identify these segments we can use cluster analysis and discover which are the tightest clusters - these would presumably be highly competitive zero-sum markets.

```{r coffee,message=FALSE}
library(proxy)
library(pvclust)
library(useful)
library(ggfortify)

coffee<-read.table("COFFEE.txt",header=FALSE)
row.names(coffee)<-coffee$V1
coffee <- coffee[,2:12]
coffee<-as.matrix(coffee)
colnames(coffee)<-row.names(coffee)
coffee_t<-t(coffee)
coffee_inv<-1-cor(coffee_t)
coffee_inv_df<-as.data.frame(coffee_inv)
coffee_dist<-as.dist(coffee_inv)
```

## Heatmap of coffee dissimilarity matrix
```{r dist}
heatmap(coffee_inv)
```

## Hierarchical clustering with `pvclust`
Pvclust uses bootstrap resampling to estimate the probability a cluster is real. This can help us determine the number of clusters to request for a k-means clustering.

```{r pvclust,message=FALSE}
#needs dissimilarity matrix
coffeeclust<-pvclust(cor(coffee_t),method.dist="correlation", method.hclust="ward", nboot=1000)
plot(coffeeclust)
```

```{r choosek}
# i am going with k of ?
k<-5
```

## K-means clustering
I went with k of `r k`
```{r kmeans}
kmres<-kmeans(coffee_dist,k)
pk<-plot.kmeans(kmres,data=coffee_inv)
```

```{r decomp}
mds<-as.data.frame(cmdscale(coffee_dist))
colnames(mds)<-c("C1","C2")
pca<-prcomp(coffee_inv)
pca.results <- data.frame(kmres$cluster, names(kmres$cluster), pca$rotation)
mds.results <- data.frame(kmres$cluster, names(kmres$cluster), mds)
ggplot(pca.results,aes(PC1,PC2,label=names.kmres.cluster.))+geom_point(aes(color=as.character(kmres.cluster)))+geom_label(nudge_x=.05,nudge_y=.04)
#ggplot(mds.results,aes(C1,C2,label=names.kmres.cluster.))+geom_point(aes(color=as.factor(kmres.cluster)))+geom_label(nudge_x=.1,nudge_y=.03)
```

The level of competition within a cluster is high (even if the coffees are from the same brand).

The within cluster sum of squares is a measure of the local intensity (smaller = more intense).

```{r pressure, echo=FALSE}
clusters<-data.frame(cluster=1:k,ss=kmres$withinss)
clusters$members<-sapply(1:k,function(x){paste(names(which(kmres$cluster==x)),collapse = ", ")})
knitr::kable(clusters)
```

It appears clusters `r clusters$cluster[order(clusters$ss)[1]]` and `r clusters$cluster[order(clusters$ss)[2]]` are particularly competitive